---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  message = FALSE
)
```

**mpmm** - fit movement persistence mixed-effects models to animal tracking data 

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)

master branch:  
[![R-CMD-check](https://github.com/ianjonsen/mpmm/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/ianjonsen/mpmm/actions/workflows/R-CMD-check.yaml)

dev branch:  

<!-- badges: end -->


`mpmm` is an R package that fits movement persistence mixed-effect models to animal tracking data for inference of linear relationships with covariates, accounting for individual variability (Jonsen et al. 2019. Ecology 100:e02566). Random effects are assumed to be approximately normal. It is assumed that the location data are either relatively error-free (e.g., GPS locations) or filtered estimates from a state-space model fitted to error-prone data (e.g., Argos locations). Models are specified using standard mixed-model formulas, as you would in `lme4` or `glmmTMB`. The underlying code for specifying and estimating fixed and random effects borrows heavily on `glmmTMB` code, but is implemented in a more limited manner in `mpmm`. Currently, for example, only diagonal or unstructured covariances are possible; interaction terms are not possible; grouping term is always assumed to be the individual animal `id` (or individual sub-tracks `tid`). 

## Installation 
First, ensure you have R version >= 3.6.0 installed (preferably R 4.0.0 or higher):

```{r get-version, eval = FALSE}
R.Version()
```

### From GitHub (source)  

On PC's running Windows, ensure you have installed [Rtools](https://cran.r-project.org/bin/windows/Rtools/) 

On Mac's, ensure you have installed the [Command Line Tools for Xcode](https://developer.apple.com/download/more/) by executing `xcode-select --install` in the terminal; or you can download the latest version from the URL (free developer registration may be required). A full Xcode install uses up a lot of disk space and is not required.

Currently, `mpmm` can only be installed from GitHub:
```{r gh-installation, eval = FALSE}
remotes::install_github("ianjonsen/mpmm")
```

Note: there can be issues getting compilers to work properly, especially on a Mac with OS X 10.13.x or higher. If you encounter install and compile issues, I recommend you consult the excellent information on the [glmmTMB](https://github.com/glmmTMB/glmmTMB) GitHub.

## Basic example

`mpmm` fits mixed models and facilitates model selection, validation and visualisation of estimated covariate relationships:
```{r example, warning=FALSE, message=FALSE, fig.width=8, fig.height=6}
library(mpmm)

fit <-
  mpmm(
    ~ ice + sal_diff + (ice |
               id),
    data = ellie.ice,
    control = mpmm_control(
      REML = TRUE,
      verbose = 0 # turn off parameter trace for tidy output
    )
  ) 

summary(fit)

plot(fit)
```
